#440 Reproducibility and Statistics Assignment Homework
TEST


<!-- This is an Rmd comment. -->
<!--
You can make comments several lines long by inserting
the text in between the the start and the end of the 
comment.

Comments are where I'll tell you what to do. You should
use the space directly below the comments to do what
the comment asks. You should also describe what you're
doing to make it obvious to the reader so they can
easily follow your logic in a fully reproducible way.
For any statistical tests you run, you should explain
the results in plain English and present the statistics
as you would in an APA style manuscript. For any symbols
you don't know how to use/can't use, feel free to use the name
instead (e.g. alpha)
--> 

<!-- Load the politics.csv data file. -->
First we load the data

```{r}
politics<-read.csv("politics.csv")
```
<!-- 
Check the structure.
subject, party, testtime, minwage, and sex need to be factors.
optimismscore and income should be integer and numeric vectors.
Fix any variables that aren't right.
-->
I am going to check the structure and change subject to a factor. Additionally I am going to change the order of "pre" and "post" so that "pre" comes first. This will make my graphs read more logically.
```{r}
str(politics)
politics$subject<-factor(politics$subject)
str(politics)
```
```{r}
politics$testtime<-factor(politics$testtime, levels=c("pre", "post"))
str(politics)
```
<!-- Find the minimum, mean, and variance for posttest
optimism scores. -->
Here I am going to find the mimimum, mean, and variance for posttest optimism scores. 
```{r}
optimismscore<-politics$optimismscore[trues]
summary(politics$optimismscore[politics$testtime=="post"])
var(politics$optimismscore[politics$testtime=="post"])
```

<!-- Create a histogram of posttest optimism scores.
Feel free to use the default ugly settings since
we didn't cover how to do this in ggplot2. If
you're a real go-getter, see if you can figure
out how to make a prettier histogram with better
labels, etc. using ggplot2. -->
Now I am going to create a histogram of the posttest optimism scores. Because each optimism score is accounted for twice, I will only assess the "post" scores in order to avoid over counting. 
```{r}
hist(politics$optimismscore[politics$testtime=="post"])
```

<!-- Run a test to determine whether political affiliation
is independent of one's gender.
Present the data supporting your analyses in graphical or
tabular form.
-->
Now I am going to run a test to demonstrate if there is a relationship between party and gender. I am going to do this by running a chi-Square Test. First, I am going to split the data in half because each observation is counted twice in a pre and post condition. 
```{r}
tab<-table(politics$party[politics$testtime=="pre"],politics$sex[politics$testtime=="pre"])
tab
chisq.test(politics$party[politics$testtime=="pre"],politics$sex[politics$testtime=="pre"])
```
This test shows no relationship between party and gender, Chi-Squared(2) = .73, p=.7
<!-- Run a test to determine whether or not males and females
have different incomes.
Create a graph to display your data that includes the relevant
means and standard errors.-->

Now I am going to conduct a t-test in order to see if males and females have significantly different incomes. The results will be demonstrated in a graph.
To create the figure demonstrating gender and income, I need to include the dplyr library. This will be used to summarize the politics data set and find the means and standard errors for the different conditions. Then I will use the ggplot2 library to create a plot. This will be done for future graphs as well.
```{r}
t.test(politics$income[politics$sex=="male" & politics$testtime=="pre"],politics$income[politics$sex=="female" & politics$testtime=="pre"])
library("dplyr")
temp<-politics%>%group_by(sex)%>%summarize(means=mean(income),
              sems=sd(income)/sqrt(length(income)))
library("ggplot2")
f<-ggplot(temp,aes(x=factor(sex),y=means))+
    geom_bar(stat="identity", color="black",fill=c("deeppink","deepskyblue2"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f
```
<!-- See if male and females with different political affiliations
have different posttest optimism scores. Create a graph to display
your data that includes the relevant means and standard errors.-->

Now I am going to use a 2-way between-subjects ANOVA to see if party affiliation and sex predict posttest optimism scores independently or in an interaction.
```{r}
summary(aov(optimismscore~party*sex,data=politics[politics$testtime=="post",]))
```
Posttest scores are related to party affiliation, F(2, 60) = 27.063, p = .4.2e-09, but not to sex, F(1, 60) = 0.04, p = .84, or an interaction between affiliation and sex, F(2, 60) = .1.21, p = .30.
<!-- Create a line graph for the same data. With
error bars of course!-->

Now I will create a line graph which illustrates the relationship between posttest optimism scores and party affiliation and gender. I need to use the gplots library in order to create the plot, and manually adjust the fill,and use hexadecimal color codes. 
```{r}
temp<-politics[politics$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Optimism Scores")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.7,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```
<!-- Run a test to see if optimism scores pre- and post
watching videos depends on sex.-->

Now I will conduct a 2-way mixed ANOVA in order to see if optimism scores of pre and post vary upon gender.
```{r}
summary(aov(optimismscore~testtime*sex+Error(subject/testtime),data=politics))
```
 Pre and post optimism scores did not vary upon gender F(1, 64) = .05, p = .83
<!-- Run a Multiple Regression Analysis to see
whether you can predict someones posttest optimism
scores from their pretest optimism scores and party
affiliation. Create a figure that shows the distributions
of pre- and posttest optimism scores and sex and create
lines that best fit the data for both sexes regardless of
whether the lines are warranted based on the statistical
analyses-->

Now I will attempt to predict the post optimism score from the pre optimism score and party affiliation. I will do this with a multiple regression test.

```{r}
 summary(lm(optimismscore[testtime=="post"]~optimismscore[testtime=="pre"]+party[testtime=="post"],data=politics)) 
```
The linear model predicting pretest and posttest optimism from their political party does not explain any variance, R^2 = .09, F(3, 62) = 180.3, p = .2.2e-16.